Individual and Differential Harm in Redistricting

### 2022-05-09 paper/paper.Rmd
AL appendix

@@

```{r boston-gen-pred, fig.width=6.5}
#| fig.cap: |
#|      Model predictions of general election vote shares and turnout, compared to actual general election results.
#|      Red reference line indicates perfect agreement.
#|      Actual and estimated results have Spearman correlations of 0.87 (vote share) and 0.84 (turnout rate).
include_graphics("figures/boston_gen_pred.pdf")
```

>>>>

To estimate voter preferences by race and precinct in Alabama, we use 2020 U.S.
Census demographic data and election returns from seven recent statewide elections: the 2016 races for president and U.S. senator, the 2018 races for governor, attorney general, and secretary of state, and the 2020 races for president and U.S. senator.
Our ecological inference model used to produce the estimates is broadly similar to that of @king1997solution. However, it jointly estimates both turnout and Democratic support across all seven elections, rather than estimating each quantity separately for each election.

Adapting the notation from the previous appendix, for precinct $j$ in election $v$, let $\vap_j$ denote the voting-age population, $\votes_{jv}$ denote the total number of votes cast for major-party candidates, $s_{jv}$ denote the Democratic vote share, and $p_{jr}$ denote the fraction of the voting-age population belonging to racial group $r$, where here $r\in\{\text{White},\text{Black},\text{Other}\}$.
Our main parameters of interest are $\turn_{jr}$ and $\supp_{jr}$, the baseline turnout and Democratic support rate by race in the precinct.

```{r alabama-ei, fig.width=6.5}
#| fig.cap: |
#|      EI model estimates of precinct-level turnout by race (a) 
#|      and Democratic support by race (b).
#|      Points are sized in proportion to their voting-age population, 
#|      and colored according to their Black share of the voting-age population.
#|      All axes are transformed to a logit scale.
include_graphics("figures/al_ei.pdf")
```

We model the observed vote counts as
\begin{align*}
    \votes_{jv} \mid \turn_{jr}, \beta^{(t)}_v
        &\sim \Binlog\qty(\vap_j, \logit(\turn_j^\top\vb p_j) + \beta^{(t)}_v) \\
    \logit(\turn_j) \mid \mu^{(t)}, \Sigma^{(t)}
        &\sim \Norm(\mu^{(t)}, \Sigma^{(t)}) \\
    s_{jv}\votes_{jv} \mid \supp_{jr}, \turn_{jr}, \beta^{(s)}_v
        &\sim \Binlog\qty(\votes_{jv}, \logit\qty(
        \supp_j^\top \frac{\turn_j\circ\vb p_j}{\turn_j^\top\vb p_j}) + \beta^{(s)}_v) \\
    \logit(\supp_j) \mid \mu^{(s)}, \Sigma^{(s)}
        &\sim \Norm(\mu^{(s)}, \Sigma^{(t)}),
\end{align*}
where the $\vb*\beta^{(t)}$ and $\vb*\beta^{(s)}$ model election-specific logit shifts, $\Binlog$ indicates the Binomial distribution with the probability paramter specified on the logit scale, and $\circ$ indicates the elementwise (Hadamard) product.
We give $\beta^{(t)}$ an i.i.d. $\Norm(0, 0.6^2)$ prior,and $\beta^{(s)}$ an i.i.d. $\Norm(0, 0.4^2)$ prior,
and $\mu^{(t)}$ an i.i.d. $\Norm(0, 1.5^2)$ prior, and $\mu^{(s)}$ an i.i.d. $\Norm(0, 1.0^2)$ prior.
We set priors independently on the variance and correlation components of $\Sigma^{(t)}$ and $\Sigma^{(s)}$; the former received $\Gam(1.5, 1.5/0.5)$ and $\Gam(1.5, 1.5/0.25)$ priors for turnout and support, while the latter received uniform priors (i.e., LKJ priors with shape parameter 1).
Several reparametrizations were made for computational efficiency; the full Stan code for the model is available as part of the replication code.

The model was fitted with a mean-field variational approximation for computational efficiency.
Posterior medians of $\turn_{jr}$ and $\supp_{jr}$ are plotted in Figure \ref{fig:alabama-ei}.
These estimates broadly track with the findings of the voting rights expert retained in *Merrill*.

@@

```{r boston-gen-pred, fig.width=6.5}
#| fig.cap: |
#|      Model predictions of general election vote shares and turnout, 
#|      compared to actual general election results.
#|      Red reference line indicates perfect agreement.
#|      Actual and estimated results have Spearman correlations of 
#|      0.87 (vote share) and 0.84 (turnout rate).
include_graphics("figures/boston_gen_pred.pdf")
```
